BayesPy: Variational Bayesian Inference in Python

### 2015-03-10 response.tex

\documentclass{article}

\usepackage{url}

\title{Review Reply to JMLR Machine Learning Open Source Software Submission (14-324):\\
BayesPy: Variational Bayesian Inference in Python}

\author{Jaakko Luttinen\\
  \url{jaakko.luttinen@aalto.fi}\\
    Aalto University, Finland}

\begin{document}

\maketitle

\section{Responses to the Reviewers of Submission 14-324 (v1)}

I would like to thank..

added features etc

% \subsection{Decision (meta-review)}

% The blah blah

% \begin{itemize}

% \end{itemize}

\subsection{Reviewer 1}

\begin{itemize}

\item \underline{syntax}\\
  yeah yeah

\item \underline{speed comparison}\\
  ooh yeeeeah

\item \underline{supported models}\\

\item \underline{scheduling}\\

\end{itemize}

\subsection{Reviewer 2}

\begin{itemize}

\item \underline{installation instructions}\\

\item \underline{user community}\\

\item \underline{speed comparison}

\end{itemize}

\subsection{Reviewer 3}

Reviewer 3 did not point out any BLAH BLAH BLAH

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End:

### 2015-03-10 bayespy.tex
Fix punctuation
Update bibliography
Add advanced VB methods paragraph

\section{Design and Features}

@@

BayesPy is designed to be simple enough for average users but flexible and
efficient enough for expert users.  One goal is to keep the syntax easy and
intuitive to read and write by making it similar to the mathematical formulation
of the model.  Other notable features of BayesPy include easy handling of
missing values, parameter expansions \citep{Qi:2007} for speeding up inference
and monitoring of the variables by plotting the posterior distributions during
the message passing iteration.  For developers, the unit testing framework helps
in finding bugs, making changes and implementing new features in a robust
manner.

@@

    \plate {} {(yn)} {$n=1,\ldots,10$};
  \end{tikzpicture}
  \caption{Graphical model of the example problem.}
  \label{fig:graphical_model}
\end{figure}
In order to demonstrate BayesPy, this section solves an extremely simple problem
but which includes the main steps of using BayesPy.  The task is to estimate the
unknown mean and precision parameters of a Gaussian distribution given ten
observations.  Thus, the likelihood function is Gaussian:
\begin{align*}
  p(y_1,\ldots,y_{10}|\mu,\tau) &= \prod^{10}_{n=1}\mathcal{N}(y_n | \mu, \tau),
\end{align*}
where the Gaussian distribution is parameterized by precision instead of
variance.  The unknown mean $\mu$ and precision $\tau$ are given broad Gaussian
and gamma priors, respectively:
\begin{align*}
  p(\mu) &= \mathcal{N}(\mu|0, 10^{-6}) & p(\tau) &= \mathcal{G}(\tau|10^{-6},
  10^{-6})
\end{align*}
Figure~\ref{fig:graphical_model} shows the graphical model of this simple model.
The model is constructed in BayesPy by creating nodes:
\begin{lstlisting}
from bayespy.nodes import GaussianARD, Gamma
mu = GaussianARD(0, 1e-6)
tau = Gamma(1e-6, 1e-6)
\end{lstlisting}
These two nodes represent the two unknown variables that we are interested in.
They are the parameters of the Gaussian distribution of the observed variable:
\begin{lstlisting}
y = GaussianARD(mu, tau, plates=(10,))
\end{lstlisting}
The number of observations is given as plates which define the number of
repetitions.  Note that the syntax of the model construction is similar to the
mathematical formulation which defines the conditional distributions for each
variable.  Thus, it is easy to understand the model from the Python code.  After
the model has been constructed, some nodes are marked as observed and given
data:
\begin{lstlisting}
y.observe([4.5, 3.9, 6.3, 5.6, 4.9, 2.8, 7.4, 6.1, 4.8, 2.1])
\end{lstlisting}
Next, we want to find the posterior approximation for our latent variables.  We
create the VB inference engine:
\begin{lstlisting}
from bayespy.inference import VB
Q = VB(y, mu, tau)
\end{lstlisting}
The inference engine takes all the stochastic nodes of the model as input.  The
message passing algorithm can be run by calling \texttt{update} method:
\begin{lstlisting}
Q.update(repeat=100)
\end{lstlisting}
The algorithm is run for 100 iterations or until it converges.  In this case,
the algorithm converges in four iterations.  The posterior approximation can be
examined, for instance, by plotting the probability density functions of the
latent variables:
\begin{lstlisting}
import bayespy.plot as bpplt
import numpy as np
bpplt.pyplot.figure()
bpplt.pdf(Q[mu], np.linspace(0, 10, num=300))
bpplt.pyplot.figure()
bpplt.pdf(Q[tau], np.linspace(0, 2, num=300))
bpplt.pyplot.show()
\end{lstlisting}
Figure~\ref{fig:posterior} shows the resulting approximate posterior
distributions.  The module \texttt{bayespy.plot} contains functions for simple
plotting tasks, such as plotting the probability density function or the Hinton
diagram of a variable.  Note that the standard \texttt{pyplot} module of
Matplotlib is available at the \texttt{plot} module of BayesPy.
\begin{figure}
  \begin{center}
    \begin{subfigure}[b]{0.4\textwidth}
      \includegraphics[width=\textwidth]{fig_mu}
      \caption{$q(\mu)$}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
      \includegraphics[width=\textwidth]{fig_tau}
      \caption{$q(\tau)$}
    \end{subfigure}
  \end{center}
  \caption{The posterior approximations of $\mu$ and $\tau$.}
  \label{fig:posterior}
\end{figure}

>>>>

\section{Features}

@@

BayesPy is designed to be simple enough for average users but flexible and
efficient enough for expert users.  One goal is to keep the syntax easy and
intuitive to read and write by making it similar to the mathematical formulation
of the model.  Missing values are easy to handle and the variables can be
monitored by plotting the posterior distributions during the learning.  BayesPy
has also preliminary support for some advanced VB methods such as stochastic
variational inference \citep{Hoffman:2013}, deterministic annealing
\citep{Katahira:2008}, collapsed inference \citep{Hensman:2012},
Riemannian-gradient-based optimization \citep{Honkela:2010}, parameter
expansions \citep{Qi:2007} and pattern searches \citep{Honkela:2003}.  For
developers, the unit testing framework helps in finding bugs, making changes and
implementing new features in a robust manner.

@@

    \plate {} {(yn)} {$n=1,\ldots,10$};
  \end{tikzpicture}
  \caption{Graphical model of the example problem.}
  \label{fig:graphical_model}
\end{figure}
In order to demonstrate BayesPy, this section solves an extremely simple problem
but which includes the main steps of using BayesPy.  The task is to estimate the
unknown mean and precision parameters of a Gaussian distribution given ten
observations.  Thus, the likelihood function is Gaussian:
\begin{align*}
  p(y_1,\ldots,y_{10}|\mu,\tau) &= \prod^{10}_{n=1}\mathcal{N}(y_n | \mu, \tau),
\end{align*}
where the Gaussian distribution is parameterized by precision instead of
variance.  The unknown mean $\mu$ and precision $\tau$ are given broad Gaussian
and gamma priors, respectively:
\begin{align*}
  p(\mu) &= \mathcal{N}(\mu|0, 10^{-6}), & p(\tau) &= \mathcal{G}(\tau|10^{-6},
  10^{-6}).
\end{align*}
Figure~\ref{fig:graphical_model} shows the graphical model of this simple model.
The model is constructed in BayesPy by creating nodes:
\begin{lstlisting}
from bayespy.nodes import GaussianARD, Gamma
mu = GaussianARD(0, 1e-6)
tau = Gamma(1e-6, 1e-6)
\end{lstlisting}
These two nodes represent the two unknown variables that we are interested in.
They are the parameters of the Gaussian distribution of the observed variable:
\begin{lstlisting}
y = GaussianARD(mu, tau, plates=(10,))
\end{lstlisting}
The number of observations is given as plates which define the number of
repetitions.  Note that the syntax of the model construction is similar to the
mathematical formulation which defines the conditional distributions for each
variable.  Thus, it is easy to understand the model from the Python code.  After
the model has been constructed, some nodes are marked as observed and given
data:
\begin{lstlisting}
y.observe([4.5, 3.9, 6.3, 5.6, 4.9, 2.8, 7.4, 6.1, 4.8, 2.1])
\end{lstlisting}
Next, we want to find the posterior approximation for our latent variables.  We
create the VB inference engine:
\begin{lstlisting}
from bayespy.inference import VB
Q = VB(y, mu, tau)
\end{lstlisting}
The inference engine takes all the stochastic nodes of the model as input.  The
message passing algorithm can be run by calling \texttt{update} method:
\begin{lstlisting}
Q.update(repeat=100)
\end{lstlisting}
The algorithm is run for 100 iterations or until it converges.  In this case,
the algorithm converges in four iterations.  The posterior approximation can be
examined, for instance, by plotting the probability density functions of the
latent variables:
\begin{lstlisting}
import bayespy.plot as bpplt
import numpy as np
bpplt.pyplot.figure()
bpplt.pdf(Q[mu], np.linspace(0, 10, num=300))
bpplt.pyplot.figure()
bpplt.pdf(Q[tau], np.linspace(0, 2, num=300))
bpplt.pyplot.show()
\end{lstlisting}
Figure~\ref{fig:posterior} shows the resulting approximate posterior
distributions.  The module \texttt{bayespy.plot} contains functions for simple
plotting tasks, such as plotting the probability density function or the Hinton
diagram of a variable.  Note that the standard \texttt{pyplot} module of
Matplotlib is available at the \texttt{plot} module of BayesPy.
\begin{figure}
  \begin{center}
    \begin{subfigure}[b]{0.4\textwidth}
      \includegraphics[width=\textwidth]{fig_mu}
      \caption{$q(\mu)$}
    \end{subfigure}
    \begin{subfigure}[b]{0.4\textwidth}
      \includegraphics[width=\textwidth]{fig_tau}
      \caption{$q(\tau)$}
    \end{subfigure}
  \end{center}
  \caption{The posterior approximations of $\mu$ and $\tau$.}
  \label{fig:posterior}
\end{figure}
